---
title: "Regression Models"
author: "Maverix13"
date: "October 19, 2015"
output: pdf_document
---

## Introduction

The report presents an analysis in the relationship automobile transmission and miles per gallon (MPG) as output. The dataset(mtcars) used in the study was extracted from the 1974 Motor Trend US magazine. This report also quantifies the MPG difference between automatic and manual transmission.

## Analysis

*Exploratory Analysis*

The dataset mtcars is loaded for current analysis. 

```{r, echo=FALSE}
library(GGally)
library(ggplot2)
data(mtcars)
head(mtcars)
fit <- lm(mpg ~ factor(am), data = mtcars)
summary(fit)$coef
```

Table above shows an intercept estimate 17.12 interpreted as mean MPG for manual transmission and slope of 7.24 interpreted as difference between the means of manual and automatic transmission with a *p-value* of 2.85e-04 which is significant. Hence, we can reject the **null hypothesis** and further investigate the effect of other variables. Figure 1 shows a graphical depiction of above analysis.

Further, pair analysis of Figure 2 shows the correlation of variables other than am may have effect on MPG. 

*Model Selection*

Referring to Figure 2, the variables having higher correlation with MPG are cyl, disp, hp and  wt. From the same correlation data, it can also be noticed that cyl has very high correlation with disp -- signifying a redundancy in variables. Based on the argument presented, further analysis will be done using cyl, hp, wt and am variables.  


# Appendix

## Exploratory Analysis

### Figure 1

```{r, echo = FALSE}
g = ggplot(data = mtcars, aes(y = mpg, x = factor(am), fill = factor(am)))
g = g + geom_violin(colour = "black", line = "dotted")
g = g + geom_boxplot(width = 0.1)
g = g + xlab("Transmission (0 = automatic, 1 = manual)") + ylab("Miles/(US) gallon")
g = g + geom_jitter(height = 0.5, aes(colour = am))
g
```

### Figure 2

```{r, echo = FALSE}
pairs = ggpairs(mtcars,  axisLabels = "none", lower = list(continuous ="smooth" ), upper=list(params=list(size=3)), params = c(method = "loess"))
pairs
```